---
title: "Application of the Box-Cox transformation in valuation"
author: "Cyrill A. Murashev,"
date: "`r Sys.Date()`"
output: github_document
bibliography: /home/kaarlahti/TresoritDrive/Methodics/My/AI_for_valuers/Book/AI_for_valuers_book/Mathstat.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract
The ability to work with numerical data is an essential skill for an appraiser. Price or cost per object or unit (e.g. square meter or foot, lot, etc.), area, power, mileage, chronological age, and more are all measured numerically. Appraisers work with numbers all the time. Recently, there has been a trend toward more evaluations using mathematical apparatus, methods of mathematical statistics, and machine learning. Today, appraiser more often use a frequentist approach to probability rather than a Bayesian approach. In this regard, they are often faced with the following problem. Many statistical tests and methods assume a normal distribution of data. But the economic data that appraisers usually deal with are most likely to have a distribution other than normal. This poses a definite problem. The purpose of this paper is to give appraisers a tool to overcome this problem. The Box-Cox transformation in many cases allows you to transform the data in such a way that their distribution is normal. This will allow you to apply parametric methods. This paper consists of three parts. The first contains a general theoretical description of the method. In the second, we will apply the basic method implementation to the data of residential real estate in Almaty. The third part will focus on some of the more difficult practical issues, in particular, we will consider the possibility of applying a two-parameter transformation model. I used the R language to write this paper. The use of Python to perform the Box-Cox transformation is discussed in a later paper.

## General description

### The concept of Power transform
In [statistics](https://en.wikipedia.org/wiki/Statistics), a power transform is a family of functions applied to create a [monotonic transformation](https://en.wikipedia.org/wiki/Monotonic_function) of data using [power functions](https://en.wikipedia.org/wiki/Power_function). It is a [data transformation](https://en.wikipedia.org/wiki/Data_transformation_(statistics)) technique used to [stabilize variance](https://en.wikipedia.org/wiki/Variance-stabilizing_transformation), make the data more [normal distribution](https://en.wikipedia.org/wiki/Normal_distribution)-like, improve the validity of measures of association (such as the [Pearson correlation](https://en.wikipedia.org/wiki/Pearson_product-moment_correlation_coefficient) between variables), and for other data stabilization procedures.

### Definition
The power transformation is defined as a continuously varying function, with respect to the power parameter $\textstyle{\lambda}$, in a piece-wise function form that makes it continuous at the point of singularity $\textstyle{(λ = 0)}$. For data vectors $\textstyle{(y_{1}, \ldots, y_{n})}$ in which each $\textstyle{y_{i} > 0}$, the power transform is
$$
\displaystyle{\begin{equation}
y_{i}^{(\lambda)} = 
\begin{cases}
\frac{y_{i}^{\lambda}-1}{\lambda(GM(y))^{\lambda-1}}, &if\ \lambda \neq 0 \\
GM(y) \ln y_{i}, &if\ \lambda = 0 \\
\end{cases}
\end{equation}}
$$

where

$$
\displaystyle{\begin{equation}
GM(y)= \biggl(\prod_{i=1}^{n}y_{i} \biggr)^\frac{1}{n} = \sqrt[n]{y_{1},y_{2}\cdots,y_{n}}
\end{equation}}
$$

is the [geometric mean](https://en.wikipedia.org/wiki/Geometric_mean) of the observations $\textstyle{y_{1},\ldots,y_{n}}$. The case for $\textstyle{\lambda = 0}$ is the limit as $\textstyle{\lambda}$ approaches $\textstyle{0}$. To see this, note that

$$
\displaystyle{\begin{equation}
y_{i}^{\lambda} = exp(\lambda \ln(y_{i})) = 1 + \lambda \ln(y_{i}) + O((\lambda \ln(y_{i}))^{2}).
\end{equation}}
$$

Then

$$
\displaystyle{\begin{equation}
\frac{y_{i}^{\lambda}-1}{\lambda} = \ln(y_{i}+O(\lambda)),
\end{equation}}
$$

and everything but $\textstyle{\ln(y_{i})}$ becomes negligible for $\textstyle{\lambda}$ sufficiently small. 
You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Bibliography
---
nocite: |
  @Wiki:Power-transform
---
